Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-18 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
Fixes

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal upto a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

>>>>

The upper bound $(K-1) \lfloor 4(R/\gamma)^2 \rfloor$ on the expected number of
mistakes of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} is
optimal up to a constant factor, as long as the number of classes $K$ is at most
$O((R/\gamma)^2)$. This lower bound is stated as
\autoref{theorem:strongly-separable-examples-mistake-lower-bound} below. The
proof of the theorem can be found in
Appendix~\ref{section:proofs-for-stringly-separable-examples} in the
supplementary material.

### 2019-01-18 ICML-submission/from-weak-separability-to-strong-separability.tex
Fixes

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that:
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with margin $\gamma'$.

@@

Since examples $(x_1,y_1), (x_2, y_2), \dots, (x_T,y_T)$ are weakly linearly separable,
all examples from class $i$ lie in
$$
R_i^+ = \!\!\!\!\! \!\!\!\!\! \bigcap_{j \in \cbr{1,2,\dots,K} \setminus \cbr{i}} \cbr{x \in \B(0,1) ~:~ \ip{w_i^* - w_j^*}{x} \ge \gamma},
$$
and all examples from the remaining classes lie in
$$
R_i^- = \!\!\!\!\! \!\!\!\!\! \bigcup_{j \in \cbr{1,2,\dots,K} \setminus \cbr{i}} \cbr{x \in \B(0,1) ~:~ \ip{w_i^* - w_j^*}{x} \le -\gamma}.
$$
Therefore, to satisfy conditions~\eqref{equation:poly-pos-i}
and~\eqref{equation:poly-neg-i}, it suffices to construct $p_i$ such that:
\begin{align}
\label{eqn:r-plus}
x \in R_i^+ & \Rightarrow p_i(x) \ge \frac {\gamma'} 2 \\
\label{eqn:r-minus}
x \in R_i^- & \Rightarrow p_i(x) \le - \frac {\gamma'} 2
\end{align}

>>>>

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with margin $\gamma'$.

@@

Since examples $(x_1,y_1), (x_2, y_2), \dots, (x_T,y_T)$ are weakly linearly separable,
all examples from class $i$ lie in
$$
R_i^+ = \!\!\!\!\! \!\!\!\!\! \bigcap_{j \in \cbr{1,2,\dots,K} \setminus \cbr{i}} \cbr{x \in \B(0,1) ~:~ \ip{w_i^* - w_j^*}{x} \ge \gamma},
$$
and all examples from the remaining classes lie in
$$
R_i^- = \!\!\!\!\! \!\!\!\!\! \bigcup_{j \in \cbr{1,2,\dots,K} \setminus \cbr{i}} \cbr{x \in \B(0,1) ~:~ \ip{w_i^* - w_j^*}{x} \le -\gamma}.
$$
Therefore, to satisfy conditions~\eqref{equation:poly-pos-i}
and~\eqref{equation:poly-neg-i}, it suffices to construct $p_i$ such that
\begin{align}
\label{eqn:r-plus}
x \in R_i^+ \qquad & \Longrightarrow \qquad p_i(x) \ge \frac {\gamma'} 2 \; , \\
\label{eqn:r-minus}
x \in R_i^- \qquad & \Longrightarrow \qquad p_i(x) \le - \frac {\gamma'} 2 \; .
\end{align}

### 2019-01-18 ICML-submission/introduction.tex
Fixes

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between the learner and the adversary. At each
timestep $t$, the adversary chooses a labeled example $(x_t, y_t)$ where the
feature vector $x_t \in \R^d$ is revealed to the learner and the correct label
$y_t$ is kept secret by the adversary. Upon receiving the feature vector $x_t$,
the learner is asked to make a prediction $\widehat{y}_t$. Immediately after
making its prediction, the learner receives feedback. In contrast with the
standard full information setting (where the feedback given is the correct label
$y_t$), here the feedback is only a binary indicator of whether the prediction
was correct or not. The protocol of the problem is formally stated below.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect: their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm, namely
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
that has an expected number of mistakes at most $O(K (R/\gamma)^2))$.
The memory complexity and per-round time complexity are both $O(dK)$.
The algorithm can be viewed as running $K$ copies of the \textsc{Binary
Perceptron} algorithm, one copy for each class. Intuitively speaking, the $O(K)$
multiplicative factor in the mistake bound is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We prove
that any (possibly randomized) algorithm must make $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the case when the examples are weakly linearly separable, we construct a
non-linear feature mapping $\phi$ and associated positive definite kernel
$k(x,x')$ that makes the examples \emph{strongly} linearly separable. We then
use the kernelized version of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} for
the strongly separable case, namely Algorithm~\ref{algorithm:kernelized}. (For
details on kernel methods see e.g.~\citet{Scholkopf-Smola-2002} or
\citet{Shawe-Taylor-Cristianini-2004}.) The kernel $k(x,x')$ corresponding to
the feature mapping $\phi$ has a simple explicit formula and can be computed in
$O(d)$ time, making Algorithm~\ref{algorithm:kernelized} computationally
efficient.

>>>>

We study the problem of \textsc{Online Multiclass Linear Classification with
Bandit Feedback}~\citep{Kakade-Shalev-Shwartz-Tewari-2008}. The problem can be
viewed as a repeated game between the learner and the adversary. At each
timestep $t$, the adversary chooses a labeled example $(x_t, y_t)$ where the
feature vector $x_t \in \R^d$ is revealed to the learner and the correct label
$y_t$ is kept secret by the adversary. Upon receiving the feature vector $x_t$,
the learner is asked to make a prediction $\widehat{y}_t$. Immediately after
making its prediction, the learner receives feedback. In contrast with the
standard full-information setting (where the feedback given is the correct label
$y_t$), here the feedback is only a binary indicator of whether the prediction
was correct or not. The protocol of the problem is formally stated below.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their proof of
Theorem 2 is incorrect: their Lemma 4 cannot be directly applied to the bandit
setting, as the algorithm only makes updates on a subset of the $w_i$'s.} We
design a simple and efficient algorithm
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
that makes at most $O(K (R/\gamma)^2))$ mistakes in expectation. Its memory
complexity and per-round time complexity are both $O(dK)$. The algorithm can be
viewed as running $K$ copies of the \textsc{Binary Perceptron} algorithm, one
copy for each class. We prove that any (possibly randomized) algorithm must make
$\Omega(K (R/\gamma)^2))$ mistakes in the worst case. The extra $O(K)$
multiplicative factor the mistake bound, as compared to the full-information
setting, is the price we pay for the bandit feedback, or more precisely, the
lack of full-information feedback.

For the case when the examples are weakly linearly separable, we construct a
non-linear feature mapping $\phi$ and associated positive definite kernel
$k(x,x')$ that makes the examples \emph{strongly} linearly separable. We then
use the kernelized version of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} for
the strongly separable case (Algorithm~\ref{algorithm:kernelized}). The kernel
$k(x,x')$ corresponding to the feature mapping $\phi$ has a simple explicit
formula and can be computed in $O(d)$ time, making
Algorithm~\ref{algorithm:kernelized} computationally efficient. For details on
kernel methods see e.g.~\citet{Scholkopf-Smola-2002} or
\citet{Shawe-Taylor-Cristianini-2004}.

### 2019-01-18 ICML-submission/notions-of-linear-separability.tex
Fixes

The notion of weak linear separability is a standard assumption used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. Namely, \citet{Crammer-Singer-2003}
prove that if the examples are weakly linearly separable with a margin $\gamma$
and the norm of the examples is bounded by $R$, then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. It is a folklore result that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any (possibly randomized) algorithm must
make $\frac{1}{2} \left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the
worst case. For completeness, we state the \textsc{Multiclass Perceptron}
algorithm and give proofs of both of these results in
Appendix~\ref{section:multiclass-perceptron-proofs} in the supplementary
material.

>>>>

The notion of weak linear separability is a standard assumption used in the
full-information setting to upper bound the number of mistakes of the
\textsc{Multiclass Perceptron} algorithm. Namely, \citet{Crammer-Singer-2003}
prove that if the examples are weakly linearly separable with a margin $\gamma$
and the norm of the examples is bounded by $R$, then \textsc{Multiclass
Perceptron} algorithm makes at most $\left\lfloor 2(R/\gamma)^2 \right \rfloor$
mistakes. It is a folklore result that \textsc{Multiclass Perceptron} is
essentially optimal in the sense that any (possibly randomized) algorithm must
make $\frac{1}{2} \left\lfloor (R/\gamma)^2 \right \rfloor$ mistakes in the
worst case. For completeness, we state the \textsc{Multiclass Perceptron}
algorithm, the upper and lower bound on the number of mistakes, and their proofs
in Appendix~\ref{section:multiclass-perceptron-proofs} in the supplementary
material.
